---
title: "S5072 Assignment2"
author: "Lan Phan"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Instructions

Your final document should be an html file knitted from an Rmarkdown file.   The html file should be uploaded to  this assignment on the course website. 


In answering each of the following questions please include a) the question as a header in your Rmarkdown report, b) then include the raw code that you used to generate any tables, and c) the top ten rows of the resulting tibble dataframe (do not include more than ten rows for any table in your report).

```{r packages}
library(dplyr)
library(tidyverse)
```

## Data Import

Use data logging NBA player season statistics to practice coding with core tidyverse functions (located in course file folder - "Season_Stats_NBA.csv".  Variable descriptions are located at the Kaggle site with the original raw data here: https://www.kaggle.com/drgilermo/nba-players-stats/data

```{r}
nba.dat <- read_csv("Seasons_Stats_NBA.csv")
nba.dat
```

# Exercise 2 Answers

Use tidyverse functions to answer the following questions:

## Q1

Which NBA player scored the most points in 1991?

Hint: Remember that tibbles don't show every column when you print them.  There are more column names below the ten printed rows.

```{r}
nba.dat %>%
  filter(Year == 1991) %>%
  arrange(desc(PTS))
```

According to the results, Michael Jordan scored the most points in 1991.

## Q2

Which player had the best free throw percentage from the year 2000 to the most recent year in the data?   If there are multiple players who have tied for the best percentage, then sort on the column with player names and select the top name in ascending order.

Hint: Some variable names in the tibble are surrounded by back ticks.  Note that when you refer to these names in your tidyverse functions you should include the back ticks.

```{r}
nba.dat %>%
  filter(Year >= 2000) %>%
  mutate(FTper = FT/FTA) %>% #R won't recognize FT. as a var in arrange
  arrange(desc(FTper), Player)
```

## Q3

Rename the variable "pos" to "position"

```{r}
nba.dat2 <- nba.dat %>%
  rename(position = Pos)

colnames(nba.dat2)
```


## Q4

Use this variable to create two variables that are called "first_position" and "second_position".  (hint:separate by splitting the position variable in two)

```{r}
nba.dat3 <- nba.dat2 %>%
  separate(position, into = c("first_position", "second_position"), sep = "-", convert = TRUE)
```


## Q5

Unite these two variables back into a single variable called "position_united"

```{r}
nba.dat3 %>%
  unite(position_united, first_position, second_position, sep = "-")
```


## Q6

Create two new datasets.  1) a new dataset from the original dataset that includes all data except the age variable (be sure to give this dataset a new name).  

```{r}
nba.dat4 <- nba.dat %>%
  select(-(Age))

nba.dat4
```

2) a new dataset from the original dataset that includes the year, the player name, and age.  

```{r}
nba.dat5 <- nba.dat %>%
  select(Year, Player, Age)

nba.dat5
```

3) add a new column to both datasets called "mergeid" that includes a sequence of numbers beginning with a 1 in the first row of the data and ending with the total number of rows in the last row of the data (hint:"df1$mergeid<-seq(1,nrow(df1),1)")

```{r}
nba.dat4$mergeid <- seq(1,nrow(nba.dat4),1)
nba.dat4

nba.dat5$mergeid <- seq(1,nrow(nba.dat5),1)  
nba.dat5

```


## Q7

Join the two datasets from question six together to recreate the original dataset plus the new merge id.

```{r}
nba.dat6 <- nba.dat4 %>%
  left_join(nba.dat5, by = "mergeid")

nba.dat6
```


## Q8

Subset the original dataset to 1995.  Group the data by year and team name and then summarise the average number of points per team. (arrange from most to least points.)

```{r}
nba.dat7 <- nba.dat %>%
  filter(Year == 1995) %>%
  group_by(Year, Tm) %>%
  summarise(mean_point = mean(PTS)) %>%
  arrange(desc(mean_point))

nba.dat7
```


## Q9

Let's reshape the data in the previous question into a wide format using tidy.  Create a wide dataset that keeps year in a single column, but spreads team names to multiple individual columns with each column delineating points per team in 1995.  (hint: you should only have one year in the resulting data)

```{r}
nba.dat8 <- nba.dat7 %>%
  spread(key = Tm, value = mean_point)

nba.dat8
```



## Q10

Now return the data to a long (tidy) format by moving teams back into a single column and points in a single column (data should include three columns: one for year, one for teams, and one for points.)

```{r}
nba.dat8 %>%
  gather(2:29, key = Team, value = points)
```

