---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

# Submission Info

**Name:** Lan Phan

**Course:** G5072 - Modern Data Structures

*Monday, July 10, 2023*

# Assignment 1

**Instructions**

Use R Notebook to create an html report that includes the following: 1. *italicized text*; 2. plot generated from R code chunk; 3. at least 2 main headers and at least one sub-header and 4. at least one link

## Project Description

The selected project and its data are posted on Tidy Tuesday on Github, a weekly social data project in R. Go [here](https://github.com/rfordatascience/tidytuesday) to learn more about *Tidy Tuesday*!

According to the description of the [Coffee Rating Data](https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-07-07/readme.md) on Github, "the data comes from Coffee Quality Database courtesy of Buzzfeed Data Scientist James LeDoux... These data were collected from the Coffee Quality Institute's review pages in January 2018."


## Random Plot

I will use the *cars* data set for convenience. 


```{r packages}
library(tidyverse)
library(dplyr)
```

The original example

```{r}
plot(cars)
```
New plot

```{r}
cars %>%
  ggplot(mapping = aes(x = dist, y = speed)) +
  geom_line(color = "gray70") + 
  geom_smooth(size = 1.1, method = "loess", se = FALSE) + #smooth the curve
  labs(x = "Distance",
       y = "Speed",
       title = "Plot from cars dataset")
```


